---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ggAU - ggplot2 themes for Aarhus University

<!-- badges: start -->
<!-- badges: end -->

The goal of ggAU is to simplify the process of creating publication-ready visualizations that follow the [Aarhus University color scheme](https://medarbejdere.au.dk/en/administration/communication/guidelines/guidelinesforcolours). 

## Installation

You can install the development version of ggAU from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("juditkisistok/ggAU")
```

## Available color palettes

```{r palettes, fig.width = 10, fig.height = 5, dpi=300}
library(ggAU)
unikn::seecol(au_color_palette(style = "light"), main = "Light", 
              grid = F, rgb = F)
unikn::seecol(au_color_palette(style = "dark"), main = "Dark", 
              grid = F, rgb = F)
unikn::seecol(au_color_palette(style = "hotandcold"), main = "Hot and cold", 
              grid = F, rgb = F)
unikn::seecol(au_color_palette(style = "hotandcold_dark"), main = "Hot and cold (dark)", 
              grid = F, rgb = F)
```

# Examples

The `scale_fill_au` and `scale_color_au` functions allow you to apply the pre-defined color palettes. 

The default color scheme is `light` and the default variable type is continuous. You can add `discrete = T` for categorical variable color schemes.

```{r scale_fill_au-examples, fig.width=8, fig.height=4, dpi = 300}
iris_df = dplyr::summarize(dplyr::group_by(iris, Species), 
                           `Mean petal width` = mean(Petal.Width))

light_plot_fill = ggplot2::ggplot(iris_df, ggplot2::aes(x = Species, 
                                      y = `Mean petal width`, fill = Species)) +
  ggplot2::geom_bar(stat = "identity") +
  ggpubr::theme_pubr() +
  scale_fill_au(discrete = T)

dark_plot_fill = ggplot2::ggplot(iris_df, ggplot2::aes(x = Species, 
                                     y = `Mean petal width`, fill = Species)) +
  ggplot2::geom_bar(stat = "identity") +
  ggpubr::theme_pubr() +
  scale_fill_au(discrete = T, style = "dark")

cowplot::plot_grid(light_plot_fill, dark_plot_fill)
```

It is also possible to define your own mix of colors - you can retrieve a vector of hex codes by color name. 

```{r scale_fill_au-custom, dpi = 300}
my_custom_style = c("blue", "yellow", "red")

ggplot2::ggplot(iris_df, ggplot2::aes(x = Species, 
                                      y = `Mean petal width`, fill = Species)) +
  ggplot2::geom_bar(stat = "identity") +
  ggpubr::theme_pubr() +
  scale_fill_au(discrete = T, style = "custom", colors = my_custom_style)
```

